---
title: 'Clasificación de uso de suelo en Zapopan: estadisticas de signaturas'
author: "Viacheslav Shalisko"
date: "16 de junio de 2017"
output: html_document
---
 
_language: output - spanish, code comments - english_

```{r setup, include=FALSE}
# set default chunk options (ouptut results and continue in case of errors)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)

# load required libraries
library(maptools)
library(raster)
library(dismo)
library(class)
```

```{r}
verbose = 0

control_parcels_path <- "Control_zones.shp"
classification_path <- "../2016_classif/random/random_points_50000_buff100_dissok.shp"

raster_path <- "../2016_corrections/compose_LC80290462016_feb_mar_8PC_incl_b8.TIF"
raster_layers_names <- c("pc1","pc2","pc3","pc4","pc5","pc6","pc7","pc8")
raster_layers_count <- 8

class_codes <- c(100,101,103,105,110,115,120,130,135,150,155,160,161,190,192,195,197,200)
class_legend <- c("Urbano",
                  "Urbano disperso",
                  "Infraestructura",
                  "SVA",
                  "Agr. riego",
                  "Agr. temporal",
                  "Pastizal",
                  "Matorral",
                  "BEsp",
                  "BTC",
                  "BTSC",
                  "BTemp disp.",
                  "BTemp dens.",
                  "Golf",
                  "AVU",
                  "VAS",
                  "BGal",
                  "Agua")
```

```{r load-spatial-data}
# load zone shape
control_parcels <- readShapePoly(control_parcels_path)
# load classification
main_class <- readShapePoly(classification_path)

# load only the first layer of the source raster
raster_l1 <- raster(raster_path)
dim(raster_l1)

# load all layers of the source raster
raster_stack<-stack(raster_path)
dim(raster_stack)
names(raster_stack) <- raster_layers_names
```

```{r draw-data, fig.height= 17, fig.width = 12}
plot(raster_l1, axes = TRUE, legend = FALSE, 
     main="Raster fuente y parcelas de control")
plot(control_parcels, border = "red", add = TRUE)
```

```{r main-sampling-folding}
# reduce number of sampling zones to less then 100 for each class
for (j in 1:length(class_codes)) {
  main_class_subset <- main_class[main_class@data$CLASS == class_codes[j],]
  subzone_count <- 1 + floor(length(main_class_subset) / 100)
  if (verbose) {
    cat("Clase: ", class_codes[j], "\n")
    cat("      Número de zonas: ", length(main_class_subset), "\n")
    cat("      Número de subzonas inicial: ", subzone_count, "\n")
  }
  main_class_kfold_index <- kfold(main_class_subset, k = subzone_count)
  main_class_subset_kfold <- main_class_subset[main_class_kfold_index == 1,]
  if (verbose) {  
    cat("      Numero de subzonas restante: ", length(main_class_subset_kfold), "\n")
  }
  if (j == 1) {
    main_class_kfold <- main_class_subset_kfold
  } else {
    main_class_kfold <- rbind(main_class_kfold, main_class_subset_kfold)
  }
}
cat("Numero de zonas de muestreo restante para todos clases: ",length(main_class_kfold), "\n")

```

```{r main-sampling, cache=TRUE}
# sampling
main_sampling <- extract(raster_stack, main_class_kfold)
```

```{r control-sampling, cache=TRUE}
# sampling
control_sampling <- extract(raster_stack, control_parcels)
```


#### Ejemplo de datos en class (promedio, sd, boxplot)

```{r}
# review example unit of data
head(main_sampling[[1]])
main_class_kfold@data$CLASS[1]
dim(main_sampling[[1]])[1]
apply(main_sampling[[1]],2,mean)
apply(main_sampling[[1]],2,sd)
boxplot(main_sampling[[1]], col = "lightgray")
```

#### Ejemplo de datos en parcela (promedio, sd, boxplot)

```{r}
# review example unit of data
head(control_sampling[[1]])
control_parcels@data$Code_1[1]
dim(control_sampling[[1]])[1]
apply(control_sampling[[1]],2,mean)
apply(control_sampling[[1]],2,sd)
boxplot(control_sampling[[1]], col = "lightgray")
```

### Tratamiento por clase
#### Reacomodo de datos por clases de muestreo principal

```{r class-rearrangement}
# rearrange data to data.frame
# create empty dataframe for main samping by class
main_sampling_df <- data.frame(matrix(ncol = raster_layers_count + 1, nrow = 0))
colnames(main_sampling_df) <- c(raster_layers_names,"CL")

# put sampling result into dataframe, the extra column is for class (CL)
for (i in 1:length(main_sampling)) {
  c_length <- dim(main_sampling[[i]])[1]
  id <- main_class_kfold@data$CLASS[i]
  fl <- cbind(main_sampling[[i]], CL = rep(id,c_length))
  main_sampling_df <- rbind(main_sampling_df,fl)
}

# optional standardize variables (only data columns)
#main_sampling_df[,c(1:raster_layers_count)] <- scale(main_sampling_df[,c(1:raster_layers_count)]) 
```

#### Resumen de datos por clase y calculo de promedios por clase

```{r class-treatment, fig.height= 6, fig.width = 12}
# visualize variables by class
boxplot( pc1 ~ CL, data = main_sampling_df, col = "lightgray")
boxplot( pc2 ~ CL, data = main_sampling_df, col = "lightgray")
boxplot( pc3 ~ CL, data = main_sampling_df, col = "lightgray")
boxplot( pc4 ~ CL, data = main_sampling_df, col = "lightgray")
boxplot( pc5 ~ CL, data = main_sampling_df, col = "lightgray")
boxplot( pc6 ~ CL, data = main_sampling_df, col = "lightgray")
boxplot( pc7 ~ CL, data = main_sampling_df, col = "lightgray")
boxplot( pc8 ~ CL, data = main_sampling_df, col = "lightgray")

# calculate class means
M1 <-aggregate( . ~ CL, data = main_sampling_df, FUN = mean)
row.names(M1) <- M1$CL
raster_layers_count1 <- raster_layers_count + 1
M1 <- M1[,c(2:raster_layers_count1)]
M1
```

#### Relación entre clases

```{r}
# Ward Hierarchical Clustering
D <- dist(M1, method = "euclidean")
fit <- hclust(D, method="ward") 
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters 
rect.hclust(fit, k=5, border="red")
```

#### Reacomodo de datos por zonas de muestreo principal

```{r zone-rearrangement}
# rearrange data to data.frame
# create empty dataframe for main samping by zone
main_sampling_df2 <- data.frame(matrix(ncol = raster_layers_count + 1, nrow = 0))
colnames(main_sampling_df2) <- c(raster_layers_names,"ZN")

# put sampling result into dataframe, the extra column is for class (ZN)
for (i in 1:length(main_sampling)) {
  c_length <- dim(main_sampling[[i]])[1]
  id <- main_class_kfold@data$DISS_CODE[i]
  fl <- cbind(main_sampling[[i]], ZN = rep(id,c_length))
  main_sampling_df2 <- rbind(main_sampling_df2,fl)
  if (verbose) {
    print(id)
    print(main_class_kfold@data$CLASS[i])
   }
}

```

#### Calculo de promedios por zona

```{r}
# calculate main sampling zone means
M3 <-aggregate( . ~ ZN, data = main_sampling_df2, FUN = mean)
row.names(M3) <- M3$ZN
raster_layers_count1 <- raster_layers_count + 1
M3 <- M3[,c(2:raster_layers_count1)]
#M3
```

### Determinar clase mas probable para cada zona de "main sampling" (K-nearets neighbour)

```{r}
# the order of elements in results of "aggregate" is not the same as in the original class,
# so it is necessary to sort the class labels in same order as training set
# following is to get original class label vector with same order as elements in M3
main_sampling_classes <- c()
M3_length <- dim(M3)[1]
for (i in 1:M3_length) {
  zone_code <- row.names(M3[i,])
  zone_subset <- main_class_kfold[main_class_kfold@data$DISS_CODE == zone_code,]
  zone_class <- zone_subset@data$CLASS
  if (verbose) {  
    cat("Zona: ", zone_code,"\n")
    cat("Clase de zona: ", zone_class, "\n")
  }
  main_sampling_classes <- c(main_sampling_classes, zone_class)
}
  
# check the dataset dimensions
#dim(main_sampling_df2[,c(1:raster_layers_count)])
#dim(M3)
#length(main_sampling_df2[,"ZN"])

# Predict the most probable class for each parcel
prediction_M3 <- knn(main_sampling_df[,c(1:8)], # training set
                  M3,                        # test set
                  main_sampling_df[,"CL"],   # training class labels
                  k = 50, prob = TRUE)

M3_predict <- cbind(M3, prediction = prediction_M3, probability = attr(prediction_M3, "prob"), original = main_sampling_classes)
table_M3_predict <- M3_predict[,c("original","prediction","probability")]
write.csv(table_M3_predict,file = "tabla_M3_predict.csv")
```

#### Tabla con valores de predicción para cada zona de "main sampling"

```{r}
table_M3_predict
```

### Tratamiento por zonas de muestreo (test)

```{r}
# rearrange data to data.frame
# create empty dataframe
control_sampling_df <- data.frame(matrix(ncol = raster_layers_count + 1, nrow = 0))
colnames(control_sampling_df) <- c(raster_layers_names,"ZN")

# take every list unit and put it into dataframe, the extra column is for index
for (i in 1:length(control_sampling)) {
  c_length <- dim(control_sampling[[i]])[1]
  id <- control_parcels@data$Id[i]
  fl <- cbind(control_sampling[[i]], ZN = rep(id,c_length))
  control_sampling_df <- rbind(control_sampling_df,fl)
}

# resumen of data per parcel
for (i in 1:length(control_sampling)) {
  if (verbose) {
    # get parcel id
    parcel_id <- control_parcels@data$Id[i]
    cat("Parcela de control: ", as.character(parcel_id),"\n")
    cat("Clase original: ",control_parcels@data$Code_2[i],"\n")      
    # subset dataframe on parcel id
    df_selection <- control_sampling_df[control_sampling_df$ZN == parcel_id,c(1:8)]    
    # get summary
    cat("Resumen","\n")
    #print(summary(df_selection))
    boxplot(df_selection, col = "lightgray")
    cat("\n\n")
  }  
}

# optional standardize variables (only data columns)
# control_sampling_df[,c(1:raster_layers_count)] <- scale(control_sampling_df[,c(1:raster_layers_count)]) 

M2 <-aggregate( . ~ ZN, data = control_sampling_df, FUN = mean)
row.names(M2) <- M2$ZN
raster_layers_count1 <- raster_layers_count + 1
M2 <- M2[,c(2:raster_layers_count1)]
M2

```

### Determinar clase mas probable para cada zona de muestreo complementario (K-nearets neighbour)

```{r}
# the order of elements in results of "aggregate" is not the same as in the original class,
# so it is necessary to sort the class labels in same order as training set
# following is to get original class label vector with same order as elements in M2
control_sampling_classes <- c()
M2_length <- dim(M2)[1]
for (i in 1:M2_length) {
  zone_code <- row.names(M2[i,])
  zone_subset <- control_parcels[control_parcels@data$Id == zone_code,]
  zone_class <- zone_subset@data$Code_2
  if (verbose) {  
    cat("Zona: ", zone_code,"\n")
    cat("Clase de zona: ", zone_class, "\n")
  }
  control_sampling_classes <- c(control_sampling_classes, zone_class)
}
  
# check the dataset dimensions
#dim(main_sampling_df[,c(1:raster_layers_count)])
#dim(M2)
#length(main_sampling_df[,"CL"])

# Predict the most probable class for each parcel
prediction_M2 <- knn(main_sampling_df[,c(1:8)], # training set
                  M2,                        # test set
                  main_sampling_df[,"CL"],   # training class labels
                  k = 50, prob = TRUE)

M2_predict <- cbind(M2, prediction = prediction_M2, probability = attr(prediction_M2, "prob"), original = control_sampling_classes)
table_M2_predict <- M2_predict[,c("original","prediction","probability")]
write.csv(table_M2_predict,file = "tabla_M2_predict.csv")

```

#### Tabla con valores de predicción para cada zona de muestreo complementario

```{r}
table_M2_predict
```
